---
title: "Task 2.2 Supervised preprocessing and visualization"
subtitle: "Modul 12: Application of Machine Learning in Health Care"
author: "**Author:** Markus Schwaiger"
date: "**Date:** April 8, 2024"
output:
html_document: default
---
- Load dataset diabetes.csv
- Define and apply a function for IQR based outlier detection
- Replace outliers with NA and select only complete cases (function complete.cases or drop_na if using tidyvers). Consider of the meaning of zeros. 
- [Optional: Describe single features (R function describe of package Hmisc)]
- Rank features using information gain (package FSelector)
- Create boxplot and distribution plots of one discriminating (i.e., highest IG score) and one non-discriminating feature (i.e., smallest IG score)
- Update your git-repository
```{r}
require(ggplot2)
require(dplyr)
require(Hmisc)
require(FSelector)
```

## Load dataset diabetes.csv and get an overview
```{r}
diab <- read.csv("diabetes.csv")
head(diab)
dim(diab)
sum(is.na(diab))
diab_num <- select(diab, where(is.numeric))
dim(diab_num)
```

- Dimension of full dataset: (768, 9)
- Number of missing values: 0
- Dimension of data with only numeric columns: (768, 8)

## IQR based outlier detection
```{r}
outlier_dedection <- function(x){
  q1 <- quantile(x, 0.25)
  q3 <- quantile(x, 0.75)
  iqr <- q3 - q1
  lower_bound <- q1 - 1.5*iqr
  upper_bound <- q3 + 1.5*iqr
  outlier <- x < lower_bound | x > upper_bound
  return(outlier)
}
```

## Replace outliers with NA and select only complete cases
```{r}
diab_num <- diab[, sapply(diab, is.numeric)]
outliers <- sapply(diab_num, outlier_dedection)
for (col in names(diab_num)) {
  diab[, col][outliers[, col]] <- NA
}
diab_complete_cases <- na.omit(diab)
dim(diab_complete_cases)
```
- Dimension of diab_complete_cases: (639, 9)


## Describe single features
```{r}
describe(diab_complete_cases)
```

## Supervised feature selection
```{r}
X <- diab_complete_cases[, -ncol(diab_complete_cases)]  # Features (select all rows and last column not)
y <- diab_complete_cases[, ncol(diab_complete_cases)]   # class labels (all rows and only last column)
ig <- information.gain(class ~., data = data.frame(X, class = y))
print(ig)

```

## Boxplot and distribution for the Feature preg and plas

```{r}
ggplot(diab_complete_cases, aes(x = class, y = preg)) +
  geom_boxplot() +
  labs(title = "Boxplot: preg by class", x = "class", y = "preg")
ggplot(diab_complete_cases, aes(x = preg, fill = class)) +
  geom_density(alpha = 0.5) +
  labs(title = "Distribution plot: preg by class", x = "preg", y = "density")



ggplot(diab_complete_cases, aes(x = class, y = plas)) +
  geom_boxplot() +
  labs(title = "Boxplot: plas by class", x = "class", y = "plas")
ggplot(diab_complete_cases, aes(x = plas, fill = class)) +
  geom_density(alpha = 0.5) +
  labs(title = "Distribution plot: plas by class", x = "plas", y = "density")

```

