---
title: "Task 2.1 Unsupervised preprocessing and visualization"
subtitle: "Modul 12: Application of Machine Learning in Health Care"
author: "**Author:** Markus Schwaiger"
date: "**Date:** April 7, 2024"
output:
html_document: default
---
- Load dataset food.csv
- Get an overview by analyzing dimension, number of missing values and feature scaling
- Preprocess data using z-transformation (R function scale)
- Perform a principal component analysis (R function princomp)
- Create a score plot using PC1 and PC2 (R function qplot)
- Update your git-repository
```{r}
require(ggplot2)
```

## Load dataset food.csv and get an overview
```{r}
food <- read.csv("food.csv", row.names = 1) # first column is used as row names
head(food)
dim(food)
sum(is.na(food))
summary(food)
```
- The first column (variable X) is used as row names
- The dataset contains primary numeric features
- The dataset consists of 25 rows and 9 columns (without the first column-variable X)
- There are no missing values in the dataset

## Preprocess data using z-transformation
```{r}
food_scale <- scale(food)
summary(food_scale)
```

## Perform a principal component analysis (PCA)
```{r}
pca <- princomp(food_scale)
summary(pca)
```

## Score plot the reduced dataset
- Comp.1 = PC1 and Comp.2 = PC2
- I used ggplot() instead of qplot() because of the description in the ggplot2 documentation, that qplot() is deprecated
```{r}
food_red <- pca$scores[, 1:2]
food_red <- food_red*-1
ggplot(data = food_red, aes(x = Comp.1, y = Comp.2,label =rownames(food_red))) +
  geom_point() +
  geom_text(size = 2, hjust = -0.2) +
  labs(x = "PC1", y = "PC2", title = "Score Plot using PC1 and PC2 by Countries")
```