---
title: "Task 3.1 Unsupervised Learning - Clustering"
subtitle: "Modul 12: Application of Machine Learning in Health Care"
author: "**Author:** Markus Schwaiger"
date: "**Date:** April 21, 2024"
output:
html_document: default
---
- Load dataset food.csv
- Preprocess data using z-transformation (R function scale)
- Apply KMeans for different number of clusters by varying k from 2 to 5. Calculate the Silhouette coefficient (function intCriteria of package clusterCrit) and select the final number of clusters
- Visualize the final kMeans clustering using PCA (R function princomp and qplot)
- Create a hierarchical clustering (R function hclust) and plot the dendrogram. In addition, create a heatmap with a clustering of samples and variables (R function heatmap)
- Create a density based clustering (R function dbscan of package fpc)
- Update your git-repository.
```{r}
require(clusterCrit)
require(factoextra)
require(ggplot2)
require(gplots)
require(fpc)

```

## Load and scale dataset food.csv
```{r}
food <- read.csv("../data/food.csv", row.names = 1) # first column is used as row names
food_scale <- scale(food)
```

## Perform KMeans clustering for different numbers of clusters (k = 2 to 5)
```{r}
set.seed(123)
for (k in 2:5) {
  kmc <- kmeans(food_scale, centers = k) #kmc=kMeans cluster
  ci <- intCriteria(food_scale, kmc$cluster, "silhouette") #ci=clusterindex
  sil_coeff <- ci$silhouette[1]
  cat("Silhouette coefficient für k=", k, ":\n")
  print(sil_coeff)
}
```

## Select the optimal number of clusters based on the silhouette coefficient
```{r}
optimal_k <- which.max(sil_coeff) + 1
final_kmeans <- kmeans(food_scale, centers = optimal_k)
cat("Optimal number of clusters:", optimal_k, "\n")
fviz_nbclust(food_scale, kmeans, method='silhouette')
```

## Visualize the final KMeans clustering using PCA
```{r}
pca <- princomp(food_scale)
food_red <- pca$scores[,1:2] 
food_red <- food_red * -1
fviz_cluster(final_kmeans, data =food_red ,   pointsize = 1, labelsize =8)
```

## Hierarchical clustering & Heatmap
```{r}
hclust_result <- hclust(dist(food_scale))
plot(hclust_result, 
     main = "Dendrogram of hierarchical clustering",
     xlab = "Country",
     ylab = "Distance",
     cex = 0.8,  
     hang = -1)
heatmap.2(food_scale, scale = "none", main = "Heatmap")

```

## Density based clustering
```{r}
dbc <- dbscan(food_scale, eps = 2.5, MinPts = 3)
fviz_cluster(dbc, data =food_red ,   pointsize = 1, labelsize =8)
```