---
title: "Task 4.2 Supervised Learning - Classification"
subtitle: "Modul 12: Application of Machine Learning in Health Care"
author: "**Author:** Markus Schwaiger"
date: "**Date:** May 3, 2024"
output:
html_document: default
---
- Load dataset diabetes.arff dataset (use R package foreign)
- Fit a Generalized Linear Model (use R function glm)
- Evaluate the model (R function summary and plot)
- Train the model using R function train of package caret
- Update your git-repository. 
```{r}
require(foreign)
require(caret)
```

# Load dataset diabetes.arff
```{r}
diab <- read.arff("../data/diabetes.arff")
head(diab)
dim(diab)
sum(is.na(diab))
```

# Fit and evaluate a generalized linear model
```{r}
diab_glm <- glm(class ~ . , data=diab, family = "binomial")
diab_glm
summary(diab_glm)
```

- A 1 unit increase in preg is associated with an increase in the odds ratio of class by 0.1231823.
- A 1 unit increase in plas is associated with an increase in the odds ratio of class by 0.0351637.
- A 1 unit increase in pres is associated with a decrease in the odds ratio of class by 0.0132955.
- A 1 unit increase in skin is associated with an increase in the odds ratio of class by 0.0006190.
- A 1 unit increase in insu is associated with a decrease in the odds ratio of class by 0.0011917.
- A 1 unit increase in mass is associated with an increase in the odds ratio of class by 0.0897010.
- A 1 unit increase in pedi is associated with an increase in the odds ratio of class by  0.9451797.
- A 1 unit increase in age is associated with an increase in the odds ratio of class by 0.0148690.

- Based on the p values we can see that at a 5% significance level, the variables skin, insu and age are not relevant predictors of diabetes for this sample.

# Train the model
### Model 1: All variabeles
```{r}
set.seed(123)
fitControl <- trainControl(method="cv",number =10)
glmFit <- train(class ~., data=diab, method = "glm", preProc=c("center","scale"), trControl = fitControl)
glmFit
```

### Model 2: Only significant variables
```{r}
set.seed(123)
glmFit2 <- train(class ~ preg + plas + pres + mass+ pedi,data=diab,method = "glm", preProc=c("center","scale"), trControl = fitControl)
glmFit2
```

### Comparing Model 1 vs. Model 2
```{r}
rbind(glmFit$results, glmFit2$results)
```

